---
title: "EDA - Variable Summary"
author: "Dave Hurst"
date: "October 10, 2015"
output: 
    html_document:
        css: standard.css

---

```{r libs, echo=FALSE, warning=FALSE, message=FALSE }
library(data.table)
library(Metrics)
library(dplyr)
library(tidyr)
library(ggvis)
library(xtable)
library(stringr)
library(magrittr)
```
```{r prep, echo=FALSE, cache=TRUE }

source( "../team/rain_utils.R")

tcheck(0)
source( "../team/data-prep.R")   ;tcheck()
#train <- train.sample1000  

#NA stats (hard code train to save time)
n_obs <- c(13765201L, 8022756L)  # train/test
n_na <- c( 5288235L, 3165860L)  # train/test
```
<span class="draft_warn">WIP: See dave/eda-var_summary.html for latest </span>


The kaggle rain contest data consists of separate training and test datasets with identical variables.  The training dataset contains a 'ground truth' value in the form of rain guage measurement.

*Measurement metadata:* 3 columns (integer)

* Id - Station ID for a particular hour ( anonymized ).  Observations are grouped by ID/hour 
* minutes_past - the minutes past the hour the data was measured 
* radardist_km - the distance between the rain guage and the radar associated with each observation (rounded to the nearest km)

*Observation Measurements:* 20 columns (continuous)
There are 4 primary measurement types

* Ref - Radar reflectivity in km
* RefComposite - Maximum vertical reflectivity above the guage in dBZ
* RhoHV - Correlation coefficient
* Zdr - Differential reflectivity in dB
* Kdp - Specific differential phase (deg/km)

Each primary measure has 4 values and is indicated by a suffix

* Nominal (no suffix)
* _5x5_10th - 10th percentile measurements in 5x5 neighborhood around the guage.
* _5x5_50th - 50th percentile
* _5x5_90th - 90th percentile

Both datasets contain a large number of observations with no valid values for any of the 20 measurements.  
```{r echo=FALSE, results='asis'}
na_table <- data.frame( dataset = c("train", "test")
                        , '_hash__of_observations' = n_obs
                        , '_hash__w_slash_o_any_values' = n_na
                        , '_percent__w_slash_o_any_values' = sprintf("%4.1f%%", 100*n_na/n_obs ))
na_table %>% xtable( ) %>% print( type='html', sanitize.colnames.function=transform.header, include.rownames = FALSE)

```

The following table summarizes the measurement variables in the training dataset.  Observations with no valid values have been removed.

```{r echo=FALSE, results='asis', cache=TRUE }

vs <- data.frame() %>%
    rbind(  train[, lapply(.SD, function(x) 100 * sum(is.na(x)) / nrow(train) )]) %>%
    rbind(  train[, lapply(.SD, min, na.rm=TRUE)]) %>%
    rbind(  train[, lapply(.SD, mean, na.rm=TRUE)]) %>%
    rbind(  train[, lapply(.SD, max, na.rm=TRUE)]) 
vs <- as.data.frame(vs)[, -c(1,2,3,24)]
tcheck()

cat('<br><span class="draft_warn">These values are currently for the training sample </span><br>')

vs_table <- data.frame( Variable = colnames(vs)
                        , '_percent__NA_Values' =      as.numeric(vs[1,]) 
                        , 'Min_Value' =  as.numeric(vs[2,])
                        , 'Mean' =           as.numeric(vs[3,])
                        , 'Max_Value' =  as.numeric(vs[4,]) )
vs_table %>% xtable( ) %>% print( type='html', sanitize.colnames.function=transform.header, include.rownames = FALSE)
tcheck()
tcheck(2)

```



