---
title: "Kaggle Rain Training Set - Variable Summary"
author: "Scott Mark"
date: "October 7, 2015"
output: pdf_document
classoption: landscape
---

```{r, echo=FALSE}
library(lattice)
#load and prep the data if this hasn't been done
source("../team/data-prep.R", echo=FALSE, print.eval=FALSE)
```
Summary of the Kaggle Rain *training* data set, which has `r nrow(train)` rows and `r ncol(train)` columns.

There are `r sum(!complete.cases(train))` rows with at least one NA, which is `r (sum(!complete.cases(train))/nrow(train))*100`% of the data set.

There are `r length(unique(train$Id))` unique Ids in the data set, representing an hourly rain gauge measurement.

**Variable Summary**

```{r, echo=FALSE}

my.summary <- t(data.frame(lapply( train[,] , function(x) rbind( 
  nmiss = sum(is.na(x)),
  min = round(min(x, na.rm=TRUE),2),
  mean = round(mean(x,na.rm=TRUE),2),
  max = round(max(x, na.rm=TRUE),2),
  P01 = round(quantile(x,.01,na.rm=TRUE,names=FALSE),3),
  P05 = round(quantile(x,.05,na.rm=TRUE,names=FALSE),3),
  P25 = round(quantile(x,.25,na.rm=TRUE,names=FALSE),3),
  P50 = round(quantile(x,.50,na.rm=TRUE,names=FALSE),3),
  P75 = round(quantile(x,.75,na.rm=TRUE,names=FALSE),3),
  P95 = round(quantile(x,.95,na.rm=TRUE,names=FALSE),3),
  P99 = round(quantile(x,.99,na.rm=TRUE,names=FALSE),3),
  variance = round(var(x, na.rm=TRUE),2),
  stdev = round(sd(x,na.rm=TRUE),2)))))

my.summary

dp <- function(x)
{
  densityplot(~x,data=train, plot.points = FALSE,
              auto.key= list(space="right",title="Density"),xlab=substitute(x))
}

```

## minutes_past
```{r, echo=FALSE}
dp(train$minutes_past)
```

## radardist_km
```{r, echo=FALSE}
dp(train$radardist_km)
```

## Ref_5x5_10th
```{r, echo=FALSE}
dp(train$Ref_5x5_10th)
```

## Ref_5x5_50th
```{r, echo=FALSE}
dp(train$Ref_5x5_50th)
```

## Ref_5x5_90th
```{r, echo=FALSE}
dp(train$Ref_5x5_90th)
```

## RefComposite
```{r, echo=FALSE}
dp(train$RefComposite)
```

## RefComposite_5x5_10th
```{r, echo=FALSE}
dp(train$RefComposite_5x5_10th)
```

## RefComposite_5x5_50th
```{r, echo=FALSE}
dp(train$RefComposite_5x5_50th)
```

## RefComposite_5x5_90th
```{r, echo=FALSE}
dp(train$RefComposite_5x5_90th)
```

## RhoHV
```{r, echo=FALSE}
dp(train$RhoHV)
```

## RhoHV_5x5_10th
```{r, echo=FALSE}
dp(train$RhoHV_5x5_10th)
```

## RhoHV_5x5_50th
```{r, echo=FALSE}
dp(train$RhoHV_5x5_50th)
```

## RhoHV_5x5_90th
```{r, echo=FALSE}
dp(train$RhoHV_5x5_90th)
```

## Zdr
```{r, echo=FALSE}
dp(train$Zdr)
```

## Zdr_5x5_10th
```{r, echo=FALSE}
dp(train$Zdr_5x5_10th)
```

## Zdr_5x5_50th
```{r, echo=FALSE}
dp(train$Zdr_5x5_50th)
```

## Zdr_5x5_90th
```{r, echo=FALSE}
dp(train$Zdr_5x5_90th)
```

## Kdp
```{r, echo=FALSE}
dp(train$Kdp)
```

## Kdp_5x5_10th
```{r, echo=FALSE}
dp(train$Kdp_5x5_10th)
```

## Kdp_5x5_50th
```{r, echo=FALSE}
dp(train$Kdp_5x5_50th)
```

## Kdp_5x5_90th
```{r, echo=FALSE}
dp(train$Kdp_5x5_90th)
```

## Expected
```{r, echo=FALSE}
dp(train$Expected)
```

 

