---
title: "Kaggle Rain Training Set - 'Expected' Distributions"
author: "Scott Mark"
date: "October 11, 2015"
output:
  html_document: default
classoption: landscape
---

```{r, echo=FALSE}
#   pdf_document:
#     latex_engine: xelatex
#TODO:
# - should really group these by ID first, then find unique values
# - plot histos with raw numbers and log scale (where needed)

library(ggplot2)
library(lattice)
library(dplyr)

#load and prep the data if this hasn't been done
source("../team/data-prep.R", echo=FALSE, print.eval=FALSE)

#TODO:  an error somewhere in these lines that breaks PDF generation
#  , which is `r length(unique(train.0to70$Expected))/nrow(train.0to70)`% of the data
#  , which is `r nrow(train.70to2000)/nrow(train)` of the overall training data
#  , which is `r format(nrow(train.2000to6000)/nrow(train), scientific = FALSE)` of the overall training data
# , which is `r format(nrow(train.6000plus)/nrow(train), scientific=FALSE)` of the overall training data
```

There are `r length(unique(train$Expected))` unique values for 'Expected' in the `r nrow(train)` rows of the training data set.

There are `r sum(is.na(train$Expected))` NA rows.

# Overall Distribution #

The overall distribution is heavily skewed right due to extreme outlier values.

```{r}
plot(histogram(train$Expected))
ggplot(train, aes(x=Expected)) +
  geom_density()
```


# Subset Distributions #

## 0 - 70 ##
```{r}
train.0to70 <- filter(train, Expected >=0, Expected <= 70)
```
The normally expected range is 0-70mm.  There are `r length(unique(train.0to70$Expected))` unique values for 'Expected' in the `r nrow(train.0to70)` rows of this subset.

The following plots show distributions within this range.

```{r}
plot(histogram(train.0to70$Expected))
ggplot(train.0to70, aes(x=Expected)) +
  geom_density()
```


## 70 - 2000 ##
```{r}
train.70to2000 <- filter(train, Expected >=70, Expected <= 2000)
```

The next subset range is 70-2000mm, as there is a slight peak above 2000.  There are `r length(unique(train.70to2000$Expected))` unique values for 'Expected' in the `r nrow(train.70to2000)` rows of this subset.


The following plots show distributions within this range.

```{r}
plot(histogram(train.70to2000$Expected))
ggplot(train.70to2000, aes(x=Expected)) +
  geom_density()
```

## 2000 - 6000 ##
```{r}
train.2000to6000 <- filter(train, Expected >=2000, Expected <= 6000)
```
The next subset range is 2000-6000mm.  There are `r length(unique(train.2000to6000$Expected))` unique values for 'Expected' in the `r nrow(train.2000to6000)` rows of this subset.

The following table shows the distribution of values in this range.

```{r}
table(train.2000to6000$Expected)
```

The following plots show distributions within this range.

```{r}
plot(histogram(train.2000to6000$Expected))
ggplot(filter(train, Expected >=2000, Expected <= 6000), aes(x=Expected)) +
  geom_density()
```


## Over 6000 ##
```{r}
train.6000plus <- filter(train, Expected >= 6000)
```

The next subset range is above 6000.  There are `r length(unique(train.6000plus$Expected))` unique values for 'Expected' in the `r nrow(train.6000plus)` rows of this subset.


The following table shows the distribution of values in this range.

```{r}
table(train.6000plus$Expected)
```


The following plots show distributions within this range.

```{r}
plot(histogram(train.6000plus$Expected))
ggplot(train.6000plus, aes(x=Expected)) +
  geom_density()
```
